---
title: "Anomaly_Detection"
author: "Group B"
date: "20/02/2021"
output: pdf_document
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning =FALSE,echo = TRUE)
```

## Abstract
 


Anomaly detection or Outlier detection identifies data points, events or observations that deviate from dataset's normal behavior. Anomalous data indicate critical incidents or potential opportunities. In order to take advantage of opportunities or fix costly problems anomaly detection has to be done in real time. Unsupervised machine learning models can be used to automate anomaly detection. Unsupervised anomaly detection algorithms scores data based on intrinsic properties of the dataset. Distances and densities are used to give an estimation what is normal and what is an outlier. Anomaly detection monitor is a tool developed for an online retailer to check product quality issues like profit opportunities and sales glitches. The application is built using R and Shinyapp following CRISP-DM framework. 



## Business Case





## Objectives

Detect point anomalies from superstore dataset using K-NN and clustering methods



### Import data

```{r message=FALSE}
#load libraries
library(readxl)
library(tidyr)
library(dplyr)
library(ggplot2)
library(anomalize)
library(lemon)
library(ggsci)


```
 

```{r}
#read data from file
superstore<-read_excel("superstore.xls")
```
```{r  include=FALSE}
#intial structure
dim(superstore) # 9994* 20
str(superstore)
```



## Data Understanding

US Superstore dataset is sourced from [US uperstore dataset](https://www.kaggle.com/juhi1994/superstore) . The dataset have online orders for  Superstores in U.S. from 2014-2018. Tableau community is the owner of the dataset. The dataset has 9994 records and 21 attributes.




```{r include=FALSE}

#print dataset details
data_superstore<-as.data.frame( summary.default(superstore))
data_superstore<-data_superstore[-c(1:42),c(1,3)]
names(data_superstore)<-c("Attribute"," Data Type")
rownames(data_superstore)<-c()
data_superstore$Description<-c("row number","unique order number","order placed date","order shipping date","shipping mode of  order","unique customer id for order","name of customer","section of product","country based on order","city based on order","state based on order","pin code","region based on order","product id of  product","category of product","sub-category of product","name of product","selling price of product","order quantity","discount on product","profit from product")
knit_print.data.frame <- lemon_print

```

```{r caption ="Dataset description" ,render=lemon_print}
data_superstore
```

## Data Preparation


```{r}
#name columns
names(superstore)<-c("rowid","orderid","order_date","ship_date","ship_mode","customer_id",
                     "customer_name","segment","country","city","state","postal_code",
                     "region","product_id","category","sub_category","product_name",
                     "sales_amt","quantity","discount","profit_amt")

#drop columns with redundant information
superstore[,c("rowid","customer_name","country")]<-NULL


#convert to date 
superstore$order_date<-as.Date(superstore$order_date,format="%Y-%m-%d")
superstore$ship_date<-as.Date(superstore$ship_date,format="%Y-%m-%d")
```



## Descriptive Analysis




### Continous variables summary

```{r}
superstore %>% 
    select_if(is.numeric)%>% 
    summary()
  
```
### Profit 
```{r}
ggplot(data=superstore)+
  geom_histogram(mapping=aes(x=profit_amt),fill="green3")+
  coord_cartesian(ylim = c(0, 100))+
  labs(title=" Profit Distribution")



```

### Sales

```{r}
ggplot(data=superstore)+
  geom_histogram(mapping=aes(x=sales_amt),fill="sienna3")+
  coord_cartesian(ylim = c(0, 100))+labs(title=" Sales Distribution")

```
### Discount


```{r}

ggplot(data=superstore)+
  geom_histogram(mapping=aes(x=discount),fill="cyan3")+
  labs(title=" Discount Distribution")
```
### Sales Profit


```{r}

ggplot(data = superstore) +
  geom_point(mapping = aes(x = sales_amt, y = profit_amt),colour="limegreen")+
  labs(title=" Sales Profit Distribution")
  

```
### Profit Discount


```{r}

ggplot(data = superstore) +
  geom_point(mapping = aes(x = profit_amt, y = discount),colour="violetred3")+
  labs(title=" Profit Discount Distribution")

```



```{r}
# total product id 
count_product_id<-unique(superstore$product_id)
length(count_product_id)


```
```{r}
#total product name
count_product_name<-unique(superstore$product_name)
length(count_product_name)



```


```{r}
#product name and product id mismatch
superstore %>% 
  distinct(product_name,product_id) %>% 
  group_by(product_id) %>% 
  filter(n()>1) %>% 
  select(product_id)


```


```{r}
#total category and subcategory 

count_category<-unique(superstore$category)
length(count_category)


count_subcategory<-unique(superstore$sub_category)
length(count_subcategory)


superstore %>% 
  distinct(category,sub_category)


```
```{r}

superstore_sales<-superstore %>% 
                  select(order_date,sales_amt)


superstore_sales<-as_tibble(superstore_sales)


```

```{r}

# superstore_sales_anomalized <- superstore_sales %>%
#     time_decompose(sales_amt, merge = TRUE) %>%
#     anomalize(remainder) %>%
#     time_recompose()



```



```{r}

```



## Test & Train dataset


```{r}

```


## Model

```{r}

```


## Evaluation

```{r}

```


## Deployment

```{r}

```


## Responsible ML Framework




## Conclusion




## Bibliography





