---
title: '**Unsupervised Learning **'
author: "by Crystal (Yunan) Zhu, Anupama r.k, Queenie Tsang"
date: "18/02/2021"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library("readxl")
library("ggplot2")
```

```{r, message=FALSE, warning=FALSE}
#get the names of the columns
nms <- names(read_excel("US_Superstore_data.xls"))

#if the column name has "Date" in it, read the column as date data type, otherwise guess the type
ct <- ifelse(grepl("^Date", nms), "date", "guess")
data <- read_excel("US_Superstore_data.xls", col_types = ct)
```


To look at some basic statistics for this dataset:
```{r}
summary(data)
```
To look at the dimensions of the data:
```{r}
dim(data)
```
The dimensions of the dataset are 9994 by 21. 

Check that the Order Date and Ship Date column type is POSIXct which is a date data type:
```{r}
data
```

To check the data type for each column: 
```{r, echo=FALSE}
str(data)
```
There are 6 numeric variables, 2 date variables, and 13 character variables.

```{r}
glimpse(data)
```

See if there are duplicates in the data and extract them:
```{r}
data %>% distinct()
```

Another way to extract only the unique columns of the dataset: 
```{r}
data_unique <- unique(data)
```

```{r}
dim(data_unique)
```
The dimensions of the dataset with only the unique rows are still 9994 by 21, so it appears there are no duplicated rows in the original dataset.


## Check for missing values in data:
```{r, echo=FALSE, message=FALSE, warning=FALSE}
##to look for missing values which are coded as NA in each column:
n=c()
for (i in 1:ncol(data)) {
  n[i]=sum(is.na(data[,i]), fixed = TRUE)
}
missing_values=paste0(colnames(data),rep("-",15),n,rep(" missing values",))
#cat("The number of missing values for each variable are:")
missing_values
```
This is 1 missing value in this dataset.

```{r , echo=FALSE, message=FALSE, warning=FALSE}
#check missing values of each column
m=c()
for (i in 1:ncol(data))  {
  m[i]=sum(grepl("?",data[,i],fixed = TRUE))
}
missval=paste0(colnames(data),rep("-",15),m,rep(" missing values",))
#cat("The number of missing values for each variable are:")
missval
```



## list rows of data that have missing values
```{r}
data[!complete.cases(data),]
```
Percentage of missing values
```{r, echo=FALSE,message=FALSE,warning=FALSE}
mean(is.na(data))
```
Remove the spaces in the column names and replace with "_" to make variable names easier to handle:
```{r, echo=FALSE, warning=FALSE}

names(data) <- gsub(" ", "_", names(data))

#check column names have been changed:
colnames(data)
```

# Exploratory Data Analysis

Find the difference between Order Date and Ship Date, and store into a new column called diff_in_days:
```{r}
data$diff_in_days<- difftime(data$Ship_Date, data$Order_Date, units = c("days"))
```

Plot Sales in relation to Order Date:
```{r, echo=FALSE, warning=FALSE}
ggplot(data = data) +
  geom_point(mapping = aes(x = Order_Date, y = Sales), xlab="Order Date", ylab="Sales")

```
Plot Profit in relation to Order Date:
```{r}
ggplot(data = data) +
  geom_point(mapping = aes(x = Order_Date, y = Profit), xlab="Order Date", ylab="Profit")
```
Some outliers for certain days

```{r}
table(data$`Sub-Category`)
```
look at the time range for these transactions, ie. start date for Order_Date column:
```{r}
min(data$Order_Date)
#[1] "2014-01-03 UTC"
```

```{r}
max(data$Order_Date)
#[1] "2017-12-30 UTC"
```
Basically this dataset covers transactions ranging from 2014-01-03 to 2017-12-30.


```{r}
ggplot(data = data) +
  geom_bar(mapping = aes(x = Category))
```
Most type of products sold belong to the Office supplies category.


```{r}
ggplot(data = data) +
  geom_bar(mapping = aes(y = `Sub-Category`))
```

```{r}
ggplot(data = data, mapping = aes(x = Sales)) +
  xlim(0, 5000) +
  geom_histogram(binwidth = 5)
```
Most sales are very few items (<500).

```{r}
ggplot(data = data, mapping = aes(x = Quantity)) +
  geom_histogram(binwidth = 5)
```
```{r}
ggplot(data = data) +
  geom_histogram(mapping = aes(x = Discount), xlab="Discount")
```
Sales transactions mostly do not involve discounts.


Visualise sales transactions by state over time (order date).
```{r}
p <- ggplot(data, aes(Order_Date, Sales)) +
      geom_line(aes(group = State), colour = "grey50") +
      geom_point(aes(colour = State))
p <- p + guides(shape = guide_legend(override.aes = list(size = 0.5)),  #this is to make legend smaller
                color = guide_legend(override.aes = list(size = 2)))
p
```

How does profit change with sub-category?
```{r}
 #density plot where the count is standardized,area under each frequency is 1 
ggplot(data = data, mapping = aes(x = Sales, y = ..density..)) +   
  geom_freqpoly(mapping = aes(colour = `Sub-Category`), binwidth = 500)
```
It looks like some categories of items ie. supplies or accessories have negative sales values.


How does sales vary across sub category?
```{r}
ggplot(data = data, mapping = aes(x = Sales, y = `Sub-Category` )) +
  geom_boxplot()
```
