---
title: '**Unsupervised Learning **'
author: "by Crystal (Yunan) Zhu, Anupama r.k, Queenie Tsang"
date: "18/02/2021"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library("readxl")
```

```{r, message=FALSE, warning=FALSE}
#get the names of the columns
nms <- names(read_excel("US_Superstore_data.xls"))

#if the column name has "Date" in it, read the column as date data type, otherwise guess the type
ct <- ifelse(grepl("^Date", nms), "date", "guess")
data <- read_excel("US_Superstore_data.xls", col_types = ct)
```


To look at some basic statistics for this dataset:
```{r}
summary(data)
```
To look at the dimensions of the data:
```{r}
dim(data)
```
The dimensions of the dataset are 9994 by 21. 

Check that the Order Date and Ship Date column type is POSIXct which is a date data type:
```{r}
data
```

To check the data type for each column: 
```{r, echo=FALSE}
str(data)
```
There are 6 numeric variables, 2 date variables, and 13 character variables.


See if there are duplicates in the data and extract them:
```{r}
data %>% distinct()
```

Another way to extract only the unique columns of the dataset: 
```{r}
data_unique <- unique(data)
```

```{r}
dim(data_unique)
```
The dimensions of the dataset with only the unique rows are still 9994 by 21, so it appears there are no duplicated rows in the original dataset.