---
title: "ML 1000 Assignment 2"
author: "by Anupama r.k, Queenie Tsang, Crystal (Yunan) Zhu"
date: "21/02/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r, echo=FALSE,message=FALSE,warning=FALSE}
library(tidyverse)
library("ggplot2")
data=read.csv("US_Superstore_data.csv")
data$Order.Date=as.Date(data$Order.Date)
data$Ship.Date=as.Date(data$Ship.Date)

diff_in_days=as.numeric(data$Ship.Date-data$Order.Date)
data=cbind(data,diff_in_days)

#check dup
#data %>% distinct()
data_nodup=distinct(data,data[,1:21], keep_all=TRUE)[,-22]
#dim did not change - no dup

#check missing values
n=c()
for (i in 1:ncol(data)) {
  n[i]=sum(is.na(data[,i]))
}
missing_values=paste0(colnames(data),rep("-",ncol(data)),n,rep(" missing values",))
#cat("The number of missing values for each variable are:")
missing_values
#no missing values
data_miss=data[!complete.cases(data),]
```


Get a general idea of the data set.


```{r}
length(unique(data$Customer.ID))
#793 unique customer IDs
length(unique(data$Customer.Name))
#793 unique customer names - drop one of these two vars


length(unique(data$Order.Date))
#1237 unique order dates
length(unique(data$Ship.Date))
#1334 unique ship dates - more unique ship dates than order dates - orders made on the same day were shipped in different dates??

length(unique(data$Segment))
unique(data$Segment)
#"Consumer"    "Corporate"   "Home Office"

unique(data$Country)
#all are from US - could drop this variable due to no-variation introduced by it

length(unique(data$City))
#531 different cities

length(unique(data$State))
#49 states

length(unique(data$Postal.Code))
#631 postal code - 793 unique customer IDs - some customers live very close!

unique(data$Region)
#only 4 regions

unique(data$Category)
#only 3 categories - "Furniture" "Office Supplies" "Technology"

length(unique(data$Sub.Category))
unique(data$Sub.Category)
#17 sub-categories 

length(unique(data$Product.Name))
#1850 product names
length(unique(data$Product.ID))
#1862 product IDs - potential redundant variables!

hist(data$diff_in_days)

summary(data$Sales)
boxplot(data$Sales)
hist(data$Sales)
#a large amount of orders with very small Sales!


summary(data$Quantity)
boxplot(data$Quantity)
#not many outliers - the #of products in each order is stable?
hist(data$Quantity)
#very skewed distribution - most of the orders have small #of items

summary(data$Discount)
boxplot(data$Discount)
#a strange looking bodataplot? - median & 3rd quantile are the same (0.2) - not many orders have high discounts
hist(data$Discount)
#most of the orders were placed without any discounts or with 20% off

summary(data$Profit)
boxplot(data$Profit)
#most of the profits are outside of the box - but most of them clustered close to the box(not with so extreme values)
hist(data$Profit)
#most of the orders have profits ~1000 (or ~800?), and ~ -800
```


Remove the dot in the column names and replace with "_" to make variable names easier to handle:

```{r, echo=FALSE, warning=FALSE}
#replace . with _ in colnames
colnames(data) <- gsub("\\.", "_", colnames(data))
#check column names have been changed:
colnames(data)
```


# Exploratory Data Analysis  


Plot Sales in relation to Order Date:  


```{r, echo=FALSE, warning=FALSE}
ggplot(data = data) +
  geom_point(mapping = aes(x = Order_Date, y = Sales), xlab="Order Date", ylab="Sales")
```
Plot Profit in relation to Order Date:
```{r}
ggplot(data = data) +
  geom_point(mapping = aes(x = Order_Date, y = Profit), xlab="Order Date", ylab="Profit")
```
Some outliers for certain days


```{r}
table(data$`Sub_Category`)
```
look at the time range for these transactions, ie. start date for Order_Date column:

```{r}
summary(data$Order_Date)
#[1] min "2014-01-03", max "2017-12-30"
```
Basically this dataset covers transactions ranging from 2014-01-03 to 2017-12-30.


```{r}
ggplot(data = data) +
  geom_bar(mapping = aes(x = Category),fill="green4")
```
Most type of products sold belong to the Office supplies category.


```{r}
ggplot(data = data) +
  geom_bar(mapping = aes(y = `Sub_Category`), fill="green4")
```

```{r}
ggplot(data = data, mapping = aes(x = Sales)) +
  xlim(0, 5000) +
  geom_histogram(binwidth = 5,fill="green4")
```
Most sales are very few items (<500).

```{r}
ggplot(data = data, mapping = aes(x = Quantity)) +
  geom_histogram(binwidth = 0.5,fill="green4")
```


```{r}
ggplot(data = data) +
  geom_histogram(mapping = aes(x = Discount), 
                 binwidth = 0.05,
                 xlab="Discount",
                 fill="green4")
```
Sales transactions mostly do not involve discounts.


Visualise sales transactions by Region over time (order date).
```{r}
 ggplot(data, aes(Order_Date, Sales,color=Region)) +
      geom_line() 
```

Let's zoom in a little bit - Visualise sales transactions by Region over time (order date).

```{r}
 ggplot(data, aes(Order_Date, Sales,color=Region)) +
      geom_line() +
      ylim(0,5000)

```

How does profit change with sub-category?
```{r}
 #density plot where the count is standardized,area under each frequency is 1 
ggplot(data = data, mapping = aes(x = Sales, y = ..density..)) +   
  geom_freqpoly(mapping = aes(colour = Sub_Category), binwidth = 500)
```
It looks like some categories of items ie. supplies or accessories have negative sales values.


How does sales vary across sub category?
```{r}
ggplot(data = data, mapping = aes(x = Sales, y = `Sub_Category` )) +
  geom_boxplot()
```


<span style="color: red;"> Add Pie charts! - by sub_category, region etc </span>


<span style="color: red;"> Create a Month variable - to see the change of sales/profits by month? </span>


<span style="color: red;"> bar charts of profits/sales by region </span>


<span style="color: red;"> Output the characteristics of the orders with the highest and lowest profits/sales - e.g. what made the order? when? bought what product? in which city/state/region? Any discount? </span>


<span style="color: red;"> relationship between discount & sales, discount & profits, sales & profits, and the role of region? </span>


<span style="color: red;"> from someone's analysis - there is no significant change between the four discount categories when it comes to Sales </span>



<span style="color: red;"> sales/profits by month, rather than by date? color by region? </span>





